The problem and promise of scale in animal social multilayer networks

### 2020-02-09 paper/manuscript.Rmd
P5 move persp + some refs

---
title: The problem and promise of scale in animal social multilayer networks
author: Alec L. Robitaille^1^, Quinn M.R. Webber^2^, Julie W. Turner^1^, Eric Vander Wal^1,2^
date: ^1^Department of Biology, Memorial University of Newfoundland, Canada; ^2^Cognitive and Behavioural Ecology Interdisciplinary Program, Memorial University of Newfoundland
output:
  # github_document
  # bookdown::pdf_document2:
  #    number_sections: false
  bookdown::word_document2:
    reference_docx: template.docx
bibliography: ../references/knitcite.bib
---

@@

Our study is motivated by behavioural, landscape, and spatial ecology to build a
conceptual and analytical framework for scale-dependent multilayer networks and
we apply this framework to a case study of caribou (*Rangifer tarandus*) in
Newfoundland, Canada [@Peignier_2019]. We use this case study to address two
principal questions associated with scale-dependent multilayer networks. First,
to what degree does the social scale of relationships, spatial scale of
landscapes, and temporal scale of analysis influence our ability to interpret
complex social systems? Second, can variation in the social, spatial, and temporal
improve predictions associated with biological differences in seasonal resource
availability and social association? We conclude with perspectives for further
development and application of scale in multilayer networks with respect to phenotypes and individual fitness (Webber & Vander Wal 2018), movement ecology and collective
movement (Jolles et al. 2019), and habitat selection and space use <!-- TODO: (REF)-->,  and .

@@

The integration of behavioural and ecological theory has shown the value of considering 
the importance of various ecological disciplines to better understand and predict animal 
social networks, including demography (Shizuka and Johnson 2019), evolutionary and 
quantitative genetics (Fisher and McAdam 2017), disease ecology (Silk et al. 2017), 
and animal personality (Sih et al. 2017). While the number of studies integrating 
social network theory into a broader array of behavioural and ecological 
sub-disciplines is impressive, only recently has the role of animal space use been 
considered in describing the causes and consequences of the social environment in the 
context of (1) phenotypes and individual fitness (Webber & Vander Wal 2018), (2) 
movement ecology and collective movement (Bode et al. 2011), and (3) habitat 
selection and space use (He et al. 2019). Importantly, much of the work linking 
space use and the social environment has been theoretical and empirical studies 
are lacking (but see Peignier et al. 2019; Pinter-Wollman 2015; Strandburg-Peshkin et al. 2015; 2017). 
Here, we discuss the role mutlilayer networks within the emerging integration 
of social and spatial ecology within the behavioural ecology literature.

@@

Importantly, the key to incorporating movement in multilayer networks across
scales is the technological advancements available to overcome previous
limitations (Hughey et al. 2018). Biologging and wildlife tracking technology 
has advanced to the point where relocations can be programmed to occur as 
frequently as each second (Börger 2016), a practice which has been applied 
to questions of collective movement (e.g Strandburg-Peskin et al. 2015). 
Meanwhile, remote sensing technology, including unmanned aerial vehicles 
and satellites, can be used to track the fine-scalemovements of  
groups over large distances and time scales (Torney et al. 2019). Using 
technology to infer behaviour has therefore become standard operating procedure
for many species in many systems. Our ability to collect and process 
remotely sensed data and infer social processes based on animal movement 
provides an opportunity to use multilayer social network analysis to answer
complex questions about animal socioecology.

Given the existing framework for constructing multilayer networks from
behavioural layers (e.g. Smith-Aguilar et al. 2019), movement layers could be
integrated with interaction or association layers. Integrating movement layers
with habitat-specific social association layers could further shed light on the
role of movement as a driver of the social and spatial environments an
individual experiences (Webber and Vander Wal 2018). In our case study, we
generated multilayer networks based on habitat-specific monolayers. Coordinated
movement of dyads, or groups, could vary based on habitat, and the approach we
outline could be used to explore mechanisms linking the social and spatial
environments. For example, social processes may be an emergent property of the
landscape because animals aggregate at resources, such as waterholes
(Chamaillé-Jammes et al. 2008). Movement to, from, or within habitat patches  or
territories can also contribute to the formation of the social environment
(Spiegel et al. 2016). For example, in sleepy lizards (*Tiliqua rugosa*) males
interacted primarily at home range edges, presumably driving territorial
behaviour (Spiegel et al. 2018). Incorporating movement ecology within a
multilayer network framework is a logical next step for a burgeoning field. In
particular, when association networks are habitat-specific, we suggest the
integration of movement and space use in multilayer networks provides novel
insights into the effects of habitat configuration on the formation of the
social environment (He et al. 2019).

>>>>

---
title: The problem and promise of scale in animal social multilayer networks
author: Alec L. Robitaille^1^, Quinn M.R. Webber^2^, Julie W. Turner^1^, Eric Vander Wal^1,2^
date: ^1^Department of Biology, Memorial University of Newfoundland, Canada; ^2^Cognitive and Behavioural Ecology Interdisciplinary Program, Memorial University of Newfoundland
output:
   github_document
  # bookdown::pdf_document2:
  #    number_sections: false
  #bookdown::word_document2:
   # reference_docx: template.docx
bibliography: ../references/knitcite.bib
---

@@

Our study is motivated by behavioural, landscape, and spatial ecology to build a
conceptual and analytical framework for scale-dependent multilayer networks and
we apply this framework to a case study of caribou (*Rangifer tarandus*) in
Newfoundland, Canada [@Peignier_2019]. We use this case study to address two
principal questions associated with scale-dependent multilayer networks. First,
to what degree does the social scale of relationships, spatial scale of
landscapes, and temporal scale of analysis influence our ability to interpret
complex social systems? Second, can variation in the social, spatial, and temporal
improve predictions associated with biological differences in seasonal resource
availability and social association? We conclude with perspectives for further
development and application of scale in multilayer networks with respect to phenotypes and individual fitness [@Webber_2018], movement ecology and collective
movement (Jolles et al. 2019), and habitat selection and space use <!-- TODO: (REF)-->,  and .

@@

The integration of behavioural and ecological theory has shown the value of considering 
the importance of various ecological disciplines to better understand and predict animal 
social networks, including demography (Shizuka and Johnson 2019), evolutionary and 
quantitative genetics (Fisher and McAdam 2017), disease ecology (Silk et al. 2017), 
and animal personality (Sih et al. 2017). While the number of studies integrating 
social network theory into a broader array of behavioural and ecological 
sub-disciplines is impressive, only recently has the role of animal space use been 
considered in describing the causes and consequences of the social environment in the 
context of (1) phenotypes and individual fitness [@Webber_2018], (2) 
movement ecology and collective movement (Bode et al. 2011), and (3) habitat 
selection and space use (He et al. 2019). Importantly, much of the work linking 
space use and the social environment has been theoretical and empirical studies 
are lacking (but see Peignier et al. 2019; Pinter-Wollman 2015; Strandburg-Peshkin et al. 2015; 2017). 
Here, we discuss the role mutlilayer networks within the emerging integration 
of social and spatial ecology within the behavioural ecology literature.

@@

Importantly, the key to incorporating movement in multilayer networks across
scales is the technological advancements available to overcome previous
limitations (Hughey et al. 2018). Biologging and wildlife tracking technology 
has advanced to the point where relocations can be programmed to occur as 
frequently as each second (Börger 2016), a practice which has been applied 
to questions of collective movement (e.g Strandburg-Peskin et al. 2015). 
Meanwhile, remote sensing technology, including unmanned aerial vehicles 
and satellites, can be used to track the fine-scale movements of groups 
over large distances and time scales (Torney et al. 2019). Using 
technology to infer behaviour has therefore become standard operating procedure
for many species in many systems. Our ability to collect and process 
remotely sensed data and infer social processes based on animal movement 
provides an opportunity to use multilayer social network analysis to answer
complex questions about animal socioecology.

Given the existing framework for constructing multilayer networks from
behavioural layers (e.g. Smith-Aguilar et al. 2019), monolayer networks that
explicity quantify movement paramters could be itnegrated. For example, edges 
could represent measures of dyadic or collective movement based on dynamic 
interaction metrics (Long et al. 2014). Integrating movement layers
with habitat-specific social association layers could further shed light on the
role of movement as a driver of the social and spatial environments an
individual experiences [@Webber_2018]. In our case study, we
generated multilayer networks based on habitat-specific monolayers. Coordinated
movement of dyads, or groups, could vary based on habitat, and the approach we
outline could be used to explore mechanisms linking the social and spatial
environments. For example, social processes may be an emergent property of the
landscape because animals aggregate at resources, such as waterholes
(Chamaillé-Jammes et al. 2008). Movement to, from, or within habitat patches  or
territories can also contribute to the formation of the social environment
(Spiegel et al. 2016). For example, in sleepy lizards (*Tiliqua rugosa*) males
interacted primarily at home range edges, presumably driving territorial
behaviour (Spiegel et al. 2018). Incorporating movement ecology within a
multilayer network framework is a logical next step for a burgeoning field. In
particular, when association networks are habitat-specific, we suggest the
integration of movement and space use in multilayer networks provides novel
insights into the effects of habitat configuration on the formation of the
social environment (He et al. 2019).
